A unified pseudo-$C_\ell$ framework

### 2018-08-06 doc/Paper/main.tex
performance commented
fixes

% Add your own macros here:
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\lstset{ %
language=python,                % choose the language of the code
basicstyle=\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
frame=false,           % adds a frame around the code
tabsize=2,          % sets default tabsize to 2 spaces
captionpos=b,           % sets the caption-position to bottom
breaklines=true,        % sets automatic line breaking
breakatwhitespace=false,    % sets if automatic breaks should only happen at whitespace
escapeinside={\%*}{*)}          % if you want to add a comment within your code
}
%\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
\newcommand{\nv}{\hat{\bf n}}
\newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
\newcommand{\todo}[1]{{\bf TODO: #1}}
\newcommand{\nmt}{{\tt NaMaster}}
\newcommand{\va}{{\mathbf{a}}}
\newcommand{\vb}{{\mathbf{b}}}
\newcommand{\mC}{{\sf C}}
\newcommand{\mM}{{\sf M}}

\newcommand{\chg}[1]{{\textcolor{blue}{#1}}}
\newcommand{\as}[1]{{\textcolor{red}{[\textbf{AS}: #1]}}}
\newcommand{\dam}[1]{{\textcolor{green}{[\textbf{DA}: #1]}}}
\newcommand{\js}[1]
{{\textcolor{orange}
{[\textbf{JS}: #1]}}}

@@

\begin{abstract}
  The pseudo-$C_\ell$ provides a fast and simpler alternative to the optimal quadratic estimator to compute angular power spectra. The algorithm can be extended to deal with contaminant deprojection and $E/B$ purification, and can therefore be applied in a wide variety of scenarios of interest for current and future cosmological observations. This paper presents \nmt, a public, validated, robust, accurate and easy-to-use software package that, for the first time, provides a unified framework to compute angular cross-power spectra of any pair of spin-0 or spin-2 fields, contaminated by an arbitrary number of linear systematics and requiring $B$- or $E$-mode purification, both in the sphere or in the flat-sky approximation. We describe the mathematical background of the estimator, including all the features above, and its software implementation in \nmt. We construct a validation suite that aims to resemble the types of observations that next-generation large-scale structure and ground-based CMB experiments will face, and use it to show that the code is able to recover the input power spectra in the most complex scenarios with no detectable bias. \nmt~can be found at \url{add_url}, and is provided with comprehensive documentation and a number of code examples to maximize its usefulness in the community.
\end{abstract}

@@

  \section{Introduction}\label{sec:intro}
    Two-point functions have proven to be the most useful summary statistic for cosmology in terms of data compression, due to the near-Gaussian nature of the modes of the cosmic density field that can be easily analyzed to extract cosmological information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}. Since the only true observables in any astronomical dataset are the intensity and polarization as a function of frequency $\nu$ and sky position $\nv$, angular 2-point correlators play a particularly central role in most cosmological analyses, both in the form of configuration-space two-point correlation functions $\xi(\theta)$ and as Fourier/harmonic-space power spectra $C_\ell$ \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}. Although correlation functions are often easier to estimate from complex observations~\citep{1993ApJ...412...64L}, a far higher level of data compression and decorrelation can be obtained through $C_\ell$s. This, together with other advantages associated with power spectra (e.g. clearer separation between linear and non-linear scales, as well as Gaussian and non-Gaussian modes~\citep{2008PhRvD..77j3013H}) has motivated a large body of literature aimed at designing optimal and efficient power spectrum estimators.

@@

        The mode-coupling matrix ${\sf M}_{\ell\ell'}$ in Eq. \ref{eq:modecoup} can be computed in terms of the harmonic coefficients of the two masks and, for spin-0 and spin-2 fields, is given by:
        \begin{align}\label{eq:m00}
          &{\sf M}^{00}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}^2\,{\sf 1},\\\label{eq:m02}
          &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,{\sf 1},\\\nonumber
          &M^{0+}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0},\\\label{eq:m22}
          &{\sf M}^{22}=\left(
                        \begin{array}{cccc}
                          M^{+}_{\ell\ell'} & 0 & 0 & M^{-}_{\ell\ell'} \\
                          0 & M^{+}_{\ell\ell'} & -M^{-}_{\ell\ell'} & 0 \\
                          0 & -M^{-}_{\ell\ell'} & M^{+}_{\ell\ell'} & 0 \\
                          M^{-}_{\ell\ell'} & 0 & 0 & M^{+}_{\ell\ell'}
                        \end{array}\right),\\\nonumber
          &M^{\pm}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}^2\frac{1\pm(-1)^{\ell+\ell'+\ell''}}{2},
        \end{align}
        where ${\sf 1}$ is the unit matrix, and $P^{vw}_\ell\equiv(2\ell+1)\,{\rm PCL}_\ell(v,w)$.

@@

    \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
      In a typical data-analysis case, we deal with contaminants that pollute the measured signal. Usually, we know something about these contaminants. As an example, in case of CMB analysis, we might have have a map of a given foreground at a different frequency which can be marginalized out. But even when we do not, we often have a tracer map of a given effect, for example, the map of dust might tell us something about systematic errors related to imperfect correction to dust reddening. Even if the dust map is non-linearly and perhaps stochastically related to the amount of reddening, as long as relation is local in space, we get a linear contamination on large scales, very much like complex astrophysics of galaxy formation nevertheless reduces to linear biasing on very large scales \dam{sounds like a very strong statement that we should either prove or give references for! I think I would rather say that you have either reduced contamination enough that a first-order Taylor term should account for the rest or that, in general, you can use a Taylor expansion where each term is a contaminant map.}\js{I agree with David, I'd go with the Taylor expansion motivation}. Thus, we generically deal with a linear model of the kind:
      \begin{equation}\label{eq:aobs}
        {\bf a}^{\rm obs}(\nv)={\bf a}^v+\sum_{i=1}^{N_a}\alpha_i{\bf f}^i(\nv),
      \end{equation}
      where the observed map ${\bf a}^{\rm obs}$ is composed of a masked signal map that receives contributions from $N_a$ known contaminants ${\bf f}^i$ and where the contamination coefficients $\alpha_i$ are unknown (note that we have implicitly applied the same mask to the contaminant templates ${\bf f}^i$). The deprojection of these contaminants from the data can be done and propagated into the estimated power spectrum in a natural way in the case of quadratic estimators \citep{1992ApJ...398..169R,2004PhRvD..69l3003S,2016MNRAS.456.2095E}, and the method can be extended to the pseudo-$C_\ell$ algorithm \citep{2017MNRAS.465.1847E}. A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}. Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
      \begin{align}\label{eq:deproj_coeff}
        &\hat{\alpha}_i=\sum_iF_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv).
      \end{align}
      The clean map is then given by
      \begin{align}\label{eq:clean}
        {\bf a}^c(\nv)
        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')\\\nonumber
        &={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv'),
      \end{align}
      where there is an implicit summation over $i$ and $j$ (which we will omit from now on).
      The harmonic coefficients of ${\bf a}^c$ are therefore given by
      \begin{equation}\label{eq:clean_f}
        {\bf a}^c_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}\,F_{ij}\sum_{{\bf l}'}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}
      \end{equation}

@@

      Two further points must be addressed before moving on. First, the computation of the deprojection bias requires an estimate of the true power spectrum ${\sf C}^{ab}_\ell$ (see Eqs. \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}). In practice, this can be done through an iterative approach, starting from an initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the estimated power spectrum in the previous iteration. A sufficiently good guess will likely be enough to remove the bias in a single iteration. Alternatively, the ensemble average of Eq. \ref{eq:dpbias0} can be seen as a convolution of ${\sf C}^{ab}_\ell$, which could be treated by computing the convolution matrix as a modified version of the standard mode-coupling matrix that accounts for mode deprojection. Estimating this convolution matrix in this case becomes significantly more computationally intensive, and therefore an iterative approach is preferred. \as{I think this is not quite true. This latter way of looking at it, in effect replaces the true power spectrum with its observed quantity, which is fine in practice, but conceptually not quite right, for the same reason e.g. error on power spectrum should be based on real rather than measured power as illustrated by low CMB quadrupole. I would just remove text after ``Alternatively...''} \dam{we can do this, but I'd like to understand what the problem with this is. Since you can understand a mask as a set of deprojected contaminants (one per masked pixel), the pseudo-$C_\ell$ mode-coupling matrix is basically a flavour of the equations above for one particular set of templates. In this sense then writing the equation above as a matrix-vector product and inverting that matrix (after binning etc.) seems like the pseudo-$C_\ell$ way to go. Are you saying that you'll get more optimal results by putting in a guess for the power spectrum because that's closer to what a QML estimator would do?}

@@

    \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
      As described in \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B}, $B$-mode purification refers to the map-level removal of the contamination from $E$-modes in a the $B$-mode component of a given map caused by an incomplete sky coverage and vice-versa. The procedure is particularly useful in situations in which the $E$-mode component of the signal is significantly larger than the $B$ modes, as is the case in CMB. In this case, removing the leakage at the power spectrum level (i.e. the standard pseudo-$C_\ell$ approach) produces a sub-optimal estimator in which the variance in the $B$-mode power spectrum is dominated by the variance of the leaked $E$-modes. This section presents the purification method and the associated modifications to the pseudo-$C_\ell$ algorithm as described in the references above. In this section we describe the algorithm in terms of purifying $B$-modes from $E$-mode contamination, but we note that the reverse, i.e. purifying $E$-modes from $B$-mode contamination is conceptually exactly the same. Moreover. our discussion is also specific for spin-2 fields.

@@

        \subsubsection{Beam deconvolution}\label{sssec:maths.beam_noise.beam}
          Including the effect of an spherically symmetric instrumental beam amounts to a simple redefinition of the mode-coupling matrix
          \begin{equation}
            {\sf M}^{s_as_b}_{\ell_1\ell_2}\rightarrow{\sf M}^{s_as_b}_{\ell_1\ell_2}W_{\ell_2}^{ab},
          \end{equation}
          where $W^{ab}_\ell$ is the product of the harmonic transform of the beams for maps $a$ and $b$. Note that it is impossible to have an anisotropic beam that is uniform over the entire sphere. However, it is possible that the beam is both anisotropic and spatially varying. In this case, the pseudo-$C_\ell$ power spetrum estimation will produce biased estimates that will be suppressed by an \emph{effective} beam. Correcting for the beam in such cases can also be done analytically, but we leave it for future work. \as{Anything to add here?} \dam{maybe be a bit more precise.}

@@

         \paragraph*{Uncorrelated noise.}
           If the noise is homogeneous but uncorrelated (i.e. $\langle {\bf n}(\nv)\,{\bf n}^\dag(\nv')\rangle={\sf 1}\,\sigma_n^2(\nv)\delta(\nv,\nv')$\footnote{Here $\delta(\nv,\nv')$ is the Dirac $\delta$-function on the sphere, and $\sigma_n^2$ is the local noise variance in one steradian.}), it is also possible to find an analytical estimate of the noise bias.
           \begin{itemize}
             \item It is not difficult to prove that the pseudo-$C_\ell$ of the noise contribution is given by
                   \begin{align}
                     {\rm PCL}_\ell({\bf n},{\bf n})&={\sf 1}\,\int \frac{d\nv}{4\pi}v^2(\nv)\sigma_n^2(\nv)\\
                                                    &={\sf 1}\alpha^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
                   \end{align}
                   where, in the second line, we have assumed that we use an inverse-variance noise weighting scheme, in which the weights map is $v(\nv)=\alpha\sigma_n^{-2}(\nv)$.
             \item The additional bias from contaminant deprojection can also be computed analytically, starting from Eq. \ref{eq:dpbias0}, as:
                   \begin{align}\nonumber
                     \Delta{\sf C}^{nn}_\ell=&-2F_{ij}{\rm PCL}_\ell\left({\bf f}^i,v^2\sigma_n^2{\bf f}^j\right)\\
                     &+F_{ij}F_{rs}\left[\int d\nv\, v^2\sigma_n^2\,{\bf f}^{j\dag}{\bf f}^r\right]\,{\rm PCL}_\ell\left({\bf f}^i,{\bf f}^s\right).
                   \end{align}
           \end{itemize}

@@

    \subsection{Flat-sky pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.flat}
      Using the full-sky expression presented above and in Appendix \ref{app:shts} for the analysis of small sky patches, where the curvature of the sphere can be neglected, is numerically inefficient for two main reasons:
      \begin{enumerate}
        \item Full-sky data formats usually store information about the full sphere, even if the data being stored is limited to a small sky patch. This can lead to an inefficient use of memory and storage space, because the standard numerical implementations of spherical transforms operate on full-sky data.
        \item Spherical harmonic transforms are notoriously slower than discrete Fourier transforms (DFTs), both due to the scaling of both algorithms (${\cal O}(N_{\rm pix}^{3/2})$ for SHTs vs. ${\cal O}(N_{\rm pix}\log(N_{\rm pix}))$ for DFTs) and because by SHTs involve operations over the full sphere.
        \item Fourier transform over a flat patch offer exact quadrature, which makes them inherently more numerically stable.
      \end{enumerate}
      For this reason, \nmt\, also supports the computation of power spectra from flat-sky maps. Appendix \ref{app:flat} describes the dictionary between spin-$s$ fields and their Fourier/harmonic coefficients defined in flat and curved skies, and we present the flat-sky versions of the pseudo-$C_\ell$ methods described in the previous sections here.

@@

          \item {\bf Mask:} for the full-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}. To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask. We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but tracers the inverse of the noise variance (see top left panel of Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citet{2018PASJ...70S...4A}. The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales. As can be seen in the top-left panel of Figure~\ref{fig:lss_mocks_flat}, the mask is not binary, and its value at each pixel correspond to the pixel's fractional masked area.

>>>>

% Add your own macros here:
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{dsfont}

\lstset{ %
language=python,                % choose the language of the code
basicstyle=\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
frame=false,           % adds a frame around the code
tabsize=2,          % sets default tabsize to 2 spaces
captionpos=b,           % sets the caption-position to bottom
breaklines=true,        % sets automatic line breaking
breakatwhitespace=false,    % sets if automatic breaks should only happen at whitespace
escapeinside={\%*}{*)}          % if you want to add a comment within your code
}
%\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
\newcommand{\nv}{\hat{\bf n}}
\newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
\newcommand{\nmt}{{\tt NaMaster}}
\newcommand{\va}{{\mathbf{a}}}
\newcommand{\vb}{{\mathbf{b}}}
\newcommand{\mC}{{\sf C}}
\newcommand{\mM}{{\sf M}}
\newcommand{\mI}{{\mathds{1}}}

\usepackage{lipsum}
\usepackage{lineno}
\linenumbers
\newcommand{\chg}[1]{{\textcolor{blue}{#1}}}
\newcommand{\as}[1]{{\textcolor{red}{[\textbf{AS}: #1]}}}
\newcommand{\dam}[1]{{\textcolor{green}{[\textbf{DA}: #1]}}}
\newcommand{\js}[1]
{{\textcolor{orange}
{[\textbf{JS}: #1]}}}
\newcommand{\todo}[1]{{\bf TODO: #1}}

@@

\begin{abstract}
  The pseudo-$C_\ell$ is an algorithm for estimation of the angular
  power and cross-power spectra that is very fast and in realistic
  cases also nearly optimal. The algorithm can be extended to deal
  with contaminant deprojection and $E/B$ purification, and can
  therefore be applied in a wide variety of scenarios of interest for
  current and future cosmological observations. This paper presents
  \nmt, a public, validated, robust, accurate and easy-to-use software
  package that, for the first time, provides a unified framework to
  compute angular cross-power spectra of any pair of spin-0 or spin-2
  fields, contaminated by an arbitrary number of linear systematics
  and requiring $B$- or $E$-mode purification, both in the sphere or
  in the flat-sky approximation. We describe the mathematical
  background of the estimator, including all the features above, and
  its software implementation in \nmt. We construct a validation suite
  that aims to resemble the types of observations that next-generation
  large-scale structure and ground-based CMB experiments will face,
  and use it to show that the code is able to recover the input power
  spectra in the most complex scenarios with no detectable
  bias. \nmt~can be found at \url{add_url}, and is provided with
  comprehensive documentation and a number of code examples to
  maximize its usefulness in the community.
\end{abstract}

@@

  \section{Introduction}\label{sec:intro}
  Two-point functions have proven to be the most useful summary
  statistic for cosmology in terms of data compression, due to the
  near-Gaussian nature of the modes of the cosmic density field that
  can be easily analyzed to extract cosmological
  information~\citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,2001PhRvD..64f3001T}. Since
  the only true observables in any astronomical dataset are the
  intensity and polarization as a function of frequency $\nu$ and sky
  position $\nv$, angular 2-point correlators play a particularly
  central role in most cosmological analyses, both in the form of
  configuration-space two-point correlation functions $\xi(\theta)$
  and as Fourier-space\footnote{In this paper we use the term
    ``Fourier'' to describe both the flat space Fourier expansion as
    well as harmonic-space expansion on the sphere. We will also use
    the term ``power spectrum'' to describe both the auto power
    spectrum of a field and cross-power spectra between two fields. }
  power spectra $C_\ell$
  \citep{2012MNRAS.427.1891A,2011PhRvD..84f3505B}.  Although
  correlation functions are often easier to estimate from complex
  observations~\citep{1993ApJ...412...64L}, a Fourier space analysis
  is advantageous for many other reasons. The most important is that
  Fourier space is a natural space for working with statistics of
  translation-invariant fields. Therefore, Fourier space analysis has
  a clearer separation between linear and non-linear scales, as well
  as Gaussian and non-Gaussian
  modes~\citep{2008PhRvD..77j3013H}. Moreover, individual band-power
  measurements are typically significantly less correlated compared to
  configuration-space analysis which makes covariance matrix
  estimation easier. This has motivated a large body of literature
  aimed at designing optimal and efficient power spectrum estimators.

@@

        The mode-coupling matrix ${\sf M}_{\ell\ell'}$ in Eq. \ref{eq:modecoup} can be computed in terms of the harmonic coefficients of the two masks and, for spin-0 and spin-2 fields, is given by:
        \begin{align}\label{eq:m00}
          &{\sf M}^{00}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}^2\,\mI,\\\label{eq:m02}
          &{\sf M}^{02}_{\ell\ell'}=M^{0+}_{\ell\ell'}\,\mI,\\\nonumber
          &M^{0+}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0},\\\label{eq:m22}
          &{\sf M}^{22}=\left(
                        \begin{array}{cccc}
                          M^{+}_{\ell\ell'} & 0 & 0 & M^{-}_{\ell\ell'} \\
                          0 & M^{+}_{\ell\ell'} & -M^{-}_{\ell\ell'} & 0 \\
                          0 & -M^{-}_{\ell\ell'} & M^{+}_{\ell\ell'} & 0 \\
                          M^{-}_{\ell\ell'} & 0 & 0 & M^{+}_{\ell\ell'}
                        \end{array}\right),\\\nonumber
          &M^{\pm}_{\ell\ell'}=\frac{2\ell'+1}{4\pi}\sum_{\ell''}P^{vw}_{\ell''}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}^2\frac{1\pm(-1)^{\ell+\ell'+\ell''}}{2},
        \end{align}
        where $\mI$ is the unit matrix, and $P^{vw}_\ell\equiv(2\ell+1)\,{\rm PCL}_\ell(v,w)$.

@@

    \subsection{Contaminant deprojection}\label{ssec:maths.deproj}
    In a typical data-analysis case, we deal with contaminants that
    pollute the measured signal. Usually, we know something about
    these contaminants. As an example, in case of CMB analysis, we
    might have have a map of a given foreground at a different
    frequency which can be marginalized out. But even when we do not,
    we often have a tracer map of a given effect, for example, the map
    of dust might tell us something about systematic errors related to
    imperfect correction to dust reddening. Even if the dust map is
    non-linearly and perhaps stochastically related to the amount of
    reddening, as long as relation is local in space, we get a linear
    contamination on large scales, very much like complex astrophysics
    of galaxy formation nevertheless reduces to linear biasing on very
    large scales (see Appendix \ref{app:lincont}). Thus, we generically deal with a
    linear model of the kind:
      \begin{equation}\label{eq:aobs}
        {\bf a}^{\rm obs}(\nv)={\bf a}^v+\sum_{i=1}^{N_a}\alpha_i{\bf f}^i(\nv),
      \end{equation}
      where the observed map ${\bf a}^{\rm obs}$ is composed of a masked signal map that receives contributions from $N_a$ known contaminants ${\bf f}^i$ and where the contamination coefficients $\alpha_i$ are unknown (note that we have implicitly applied the same mask to the contaminant templates ${\bf f}^i$). The deprojection of these contaminants from the data can be done and propagated into the estimated power spectrum in a natural way in the case of quadratic estimators \citep{1992ApJ...398..169R,2004PhRvD..69l3003S,2016MNRAS.456.2095E}, and the method can be extended to the pseudo-$C_\ell$ algorithm \citep{2017MNRAS.465.1847E}. A best-fit value for the coefficients $\alpha_i$ can be found assuming \emph{equal weights} across the map\footnote{The assumption of uncorrelated weights is the key difference between pseudo-$C_\ell$ and optimal quadratic estimators \citep{2013MNRAS.435.1857L}. Note that, since the templates are already multiplied by the mask, the weights are not equal across the map, but correspond to the mask values in general.}
      \begin{align}\label{eq:deproj_coeff}
        &\hat{\alpha}_i=\sum_iF_{ij}\int d\nv\,{\bf f}^{j\dag}(\nv)\,{\bf a}^{\rm obs}(\nv),\\\nonumber
        &({\sf F}^{-1})_{ij}\equiv\int d\nv\, {\bf f}^{i\dag}(\nv)\,{\bf f}^j(\nv).
      \end{align}
      The clean map is then given by
      \begin{align}\label{eq:clean}
        {\bf a}^c(\nv)
        &\equiv{\bf a}^{\rm obs}(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^{\rm obs}(\nv')\\\nonumber
        &={\bf a}^v(\nv)-{\bf f}^i(\nv)F_{ij}\int d\nv'\,{\bf f}^{j\dag}(\nv')\,{\bf a}^v(\nv'),
      \end{align}
      where there is an implicit summation over $i$ and $j$ (which we will omit from now on).
      The harmonic coefficients of ${\bf a}^c$ are therefore given by
      \begin{equation}\label{eq:clean_f}
        {\bf a}^c_{\bf l}={\bf a}^v_{\bf l}-{\bf f}^i_{\bf l}\,F_{ij}\sum_{{\bf l}'}{\bf f}^{j\dag}_{{\bf l}'}{\bf a}^v_{{\bf l}'}
      \end{equation}

@@

      Two further points must be addressed before moving on. First,
      the computation of the deprojection bias requires an estimate of
      the true power spectrum ${\sf C}^{ab}_\ell$ (see
      Eqs. \ref{eq:dp.g.aux} and \ref{eq:dp.f.aux}). In practice, this
      can be done through an iterative approach, starting from an
      initial guess of ${\sf C}^{ab}_\ell$ and replacing it by the
      estimated power spectrum in the previous iteration. A
      sufficiently good guess will likely be enough to remove the bias
      in a single iteration.  Alternatively, the ensemble average of
      Eq. \ref{eq:dpbias0} can be seen as a convolution of
      ${\sf C}^{ab}_\ell$, which could be treated by computing the
      convolution matrix as a modified version of the standard
      mode-coupling matrix that accounts for mode
      deprojection. Estimating this convolution matrix in this case
      becomes significantly more computationally intensive, and
      therefore an iterative approach is preferred.

@@

    \subsection{\texorpdfstring{$E/B$}{E/B} purification}\label{ssec:maths.pureb}
    As described in
    \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2010PhRvD..82b3001Z,2010A&A...519A.104K,2011PhRvD..83h3003B},
    $B$-mode purification refers to the map-level removal of the
    contamination from $E$-modes in a the $B$-mode component of a
    given map caused by an incomplete sky coverage and vice-versa. The
    procedure is particularly useful in situations in which the
    $E$-mode component of the signal is significantly larger than the
    $B$ modes, as is the case in CMB. In this case, removing the
    leakage at the power spectrum level (i.e. the standard
    pseudo-$C_\ell$ approach) produces a sub-optimal estimator in
    which the variance in the $B$-mode power spectrum is dominated by
    the variance of the leaked $E$-modes. This section presents the
    purification method and the associated modifications to the
    pseudo-$C_\ell$ algorithm as described in the references above. In
    this section we describe the algorithm in terms of purifying
    $B$-modes from $E$-mode contamination, but we note that the
    reverse, i.e. purifying $E$-modes from $B$-mode contamination is
    conceptually exactly the same. Our discussion is also
    specific for the spin-2 fields.

@@

        \subsubsection{Beam deconvolution}\label{sssec:maths.beam_noise.beam}
          Including the effect of an spherically symmetric instrumental beam amounts to a simple redefinition of the mode-coupling matrix
          \begin{equation}
            {\sf M}^{s_as_b}_{\ell_1\ell_2}\rightarrow{\sf M}^{s_as_b}_{\ell_1\ell_2}W_{\ell_2}^{ab},
          \end{equation}
          where $W^{ab}_\ell$ is the product of the harmonic transform
          of the beams for maps $a$ and $b$. Note that due to hairy
          ball theorem, it is impossible to have an anisotropic beam
          that is uniform over the entire sphere.  However, it is
          possible that the beam is both anisotropic and spatially
          varying. In this case, the pseudo-$C_\ell$ power spetrum
          estimation will produce biased estimates that will be
          suppressed by an \emph{effective} beam. Correcting for the
          beam in such cases can also be done analytically, but we
          leave it for future work.  In practice, if required, the
          suppression factor can be found by applying the PCL
          estimator to a a set of mock input maps with the same
          underlying signal but  with and without application of the
          beam smoothing (to both data maps and contaminant maps) and taking
          the ration. \as{How does this interact with E/B purification?}

@@

         \paragraph*{Uncorrelated noise.}
           If the noise is homogeneous but uncorrelated (i.e. $\langle {\bf n}(\nv)\,{\bf n}^\dag(\nv')\rangle=\mI\,\sigma_n^2(\nv)\delta(\nv,\nv')$\footnote{Here $\delta(\nv,\nv')$ is the Dirac $\delta$-function on the sphere, and $\sigma_n^2$ is the local noise variance in one steradian.}), it is also possible to find an analytical estimate of the noise bias.
           \begin{itemize}
             \item It is not difficult to prove that the pseudo-$C_\ell$ of the noise contribution is given by
                   \begin{align}
                     {\rm PCL}_\ell({\bf n},{\bf n})&=\mI\,\int \frac{d\nv}{4\pi}v^2(\nv)\sigma_n^2(\nv)\\
                                                    &=\mI\alpha^2\,\int \frac{d\nv}{4\pi}\sigma_n^{-2}(\nv),
                   \end{align}
                   where, in the second line, we have assumed that we use an inverse-variance noise weighting scheme, in which the weights map is $v(\nv)=\alpha\sigma_n^{-2}(\nv)$.
             \item The additional bias from contaminant deprojection can also be computed analytically, starting from Eq. \ref{eq:dpbias0}, as:
                   \begin{align}\nonumber
                     \Delta{\sf C}^{nn}_\ell=&-2F_{ij}{\rm PCL}_\ell\left({\bf f}^i,v^2\sigma_n^2{\bf f}^j\right)\\
                     &+F_{ij}F_{rs}\left[\int d\nv\, v^2\sigma_n^2\,{\bf f}^{j\dag}{\bf f}^r\right]\,{\rm PCL}_\ell\left({\bf f}^i,{\bf f}^s\right).
                   \end{align}
           \end{itemize}

@@

    \subsection{Flat-sky pseudo-\texorpdfstring{$C_\ell$}{Cl}s}\label{ssec:maths.flat}
      Using the full-sky expression presented above and in Appendix \ref{app:shts} for the analysis of small sky patches, where the curvature of the sphere can be neglected, is numerically inefficient for two main reasons:
      \begin{enumerate}
        \item Full-sky data formats usually store information about the full sphere, even if the data being stored is limited to a small sky patch. This can lead to an inefficient use of memory and storage space, because the standard numerical implementations of spherical transforms operate on full-sky data.
        \item Spherical harmonic transforms are notoriously slower than discrete Fourier transforms (DFTs), both due to the scaling of both algorithms (${\cal O}(N_{\rm pix}^{3/2})$ for SHTs vs. ${\cal O}(N_{\rm pix}\log(N_{\rm pix}))$ for DFTs) and because by SHTs involve operations over the full sphere.
        \item Fourier transforms over a flat patch offer exact quadrature, which makes them inherently more numerically stable.
      \end{enumerate}
      For this reason, \nmt\, also supports the computation of power spectra from flat-sky maps. Appendix \ref{app:flat} describes the dictionary between spin-$s$ fields and their Fourier/harmonic coefficients defined in flat and curved skies, and we present the flat-sky versions of the pseudo-$C_\ell$ methods described in the previous sections here.

@@

          \item {\bf Mask:} for the full-sky data, we use a mask built
            as a combination of the sky coverage output by the LSST
            {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a
            more conservative galactic cut based on the reddening map
            of \citet{1998wfsc.conf..297S}. To add an extra layer of
            complexity, we drill 100 additional 1-degree holes on the
            resulting combined mask. We also explore the improvement
            stemming from the down-weighting of higher-noise in an
            inverse-variance way in the presence of inhomogeneous
            noise (see description below), which is the reason why the
            default version of this mask is not binary, but tracers
            the inverse of the noise variance (see top left panel of
            Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we
            use a high-resolution mask constructed from the
            bright-object mask used for the first public data release
            from the HSC collaboration~\citep{2018PASJ...70S...4A}. The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales. As can be seen in the top-left panel of Figure~\ref{fig:lss_mocks_flat}, the mask is not binary, and its value at each pixel correspond to the pixel's fractional masked area.
